---
title: "Diabetes Prediction"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(readr)
library(klaR) # Naïve Bayes
library(e1071) # SVM
library(ROCR) # ROC curves and AUC
library(VennDiagram)
library(RColorBrewer)
library(vioplot)
library(hfunk)
library(gplots) # heatmap.2
library(caret) # confusionmatrix
library(randomForest)

```

```{r warning=FALSE, message=FALSE}
data1 = read_csv('diabetes-dataset.csv')
data2 = read_csv('diabetes_data_upload.csv')
draw_confusion_matrix <- function(cm) {

  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)

  # create the matrix 
  rect(150, 430, 240, 370, col='#3F97D0')
  text(195, 435, 'Class1', cex=1.2)
  rect(250, 430, 340, 370, col='#F7AD50')
  text(295, 435, 'Class2', cex=1.2)
  text(125, 370, 'Predicted', cex=1.3, srt=90, font=2)
  text(245, 450, 'Actual', cex=1.3, font=2)
  rect(150, 305, 240, 365, col='#F7AD50')
  rect(250, 305, 340, 365, col='#3F97D0')
  text(140, 400, 'Class1', cex=1.2, srt=90)
  text(140, 335, 'Class2', cex=1.2, srt=90)

  # add in the cm results 
  res <- as.numeric(cm$table)
  text(195, 400, res[1], cex=1.6, font=2, col='white')
  text(195, 335, res[2], cex=1.6, font=2, col='white')
  text(295, 400, res[3], cex=1.6, font=2, col='white')
  text(295, 335, res[4], cex=1.6, font=2, col='white')

  # add in the specifics 
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)

  # add in the accuracy information 
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
  text(70, 35, names(cm$overall[2]), cex=1.5, font=2)
  text(70, 20, round(as.numeric(cm$overall[2]), 3), cex=1.4)
}  
```

Complete Model Evalutation on 2000 Patient Data
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### SVM Prediction Model Confusion Matrix
```{r}
# SVM
data1$Outcome = as.factor(data1$Outcome)
testidx <- which(1:(dim(data1[,1])[1])%%5 == 0)
dftrain <- data1[-testidx,]
dftest <- data1[testidx,]
# hist(as.numeric(dftrain$Outcome))
# hist(as.numeric(dftest$Outcome))
svm_model <- svm(Outcome~., data = dftrain, probability = T, kernel="polynomial", degree=5)
svm_prediction <- predict(svm_model, dftest, probability = T)
# table(predicted=svm_prediction, observed=dftest$Outcome)
cmsvm=confusionMatrix(svm_prediction, dftest$Outcome, positive = "1")
draw_confusion_matrix(cmsvm)
```

### SVM Prediction Model 5 Folds Cross Validation ROC
```{r}
# Repeat using 5fold cross validation. Split data into 5 equally sized parts
folds <- function(x,n) split(x, sort(rep(1:n,len=length(x))))
crossdata <- folds(data1,5)
pred_label <- list()
actual_label <- list()
auc_list <- list()

for (i in 1:5) {
  test <- do.call("rbind", crossdata[i])
  train <- do.call("rbind", crossdata[-i])
  # print((train))
  model <- svm(Outcome~., data = train, probability = T, kernel="linear")

  prediction <- predict(model, test, probability = T)
  pred_class_svm <- attr(prediction, "probabilities")[, "1"]
  table(test$Outcome, pred_class_svm)
  pred_label[[i]] <- pred_class_svm
  actual_label[[i]] <- test$Outcome == 1
  
}
cross_pred <- prediction(pred_label, actual_label)
cross_perf <- performance(cross_pred,"tpr","fpr")
svmauc_c <- performance(cross_pred, "auc")
svmauc_c <- unlist(slot(svmauc_c, "y.values"))

plot(cross_perf,col="grey82",lty=3)
legend(0.6,0.9,c(c(paste('AUC is', svmauc_c)),"\n"),
       border="white",cex=0.5, box.col = "white")
plot(cross_perf, lwd=3,avg="vertical",spread.estimate="boxplot",add=T)
```

Column {data-width=500}
-----------------------------------------------------------------------
### RF Prediction Model Confusion Matrix
```{r}
# RF
rf_classifier = randomForest(Outcome ~ ., data=dftrain, ntree=100, mtry=2, importance=TRUE, probability = T)
prediction_for_table <- predict(rf_classifier,dftest[,-9], probability = T)
# table(observed=dftest$Outcome,predicted=prediction_for_table)
cmrf=confusionMatrix(prediction_for_table, dftest$Outcome, positive = "1")
draw_confusion_matrix(cmrf)
```


### RF Prediction Model 5-Folds Cross Validation ROC
```{r}
# Repeat using 10fold cross validation. Split data into 10 equally sized parts
folds <- function(x,n) split(x, sort(rep(1:n,len=length(x))))
crossdata <- folds(data1,5)
pred_label <- list()
actual_label <- list()
auc_list <- list()

for (i in 1:5) {
  test <- do.call("rbind", crossdata[i])
  train <- do.call("rbind", crossdata[-i])
  # print((train))
  rf_classifier <- randomForest(Outcome ~ ., data=dftrain, ntree=100, mtry=2, importance=TRUE, probability = T)
  pred_class_rf <- predict(rf_classifier,test[,-9], probability = T)


  pred_prob_rf <- predict(rf_classifier,test[,-9],type="prob")[,"1"]

  pred_label[[i]] <- pred_prob_rf
  actual_label[[i]] <- test$Outcome == 1
  
}
cross_pred <- prediction(pred_label, actual_label)
cross_perf <- performance(cross_pred,"tpr","fpr")
svmauc_c <- performance(cross_pred, "auc")
svmauc_c <- unlist(slot(svmauc_c, "y.values"))

plot(cross_perf,col="grey82",lty=3)
legend(0.6,0.9,c(c(paste('AUC is', svmauc_c)),"\n"),
       border="white",cex=0.5, box.col = "white")
plot(cross_perf, lwd=3,avg="vertical",spread.estimate="boxplot",add=T)
```



Test Your Chance of Getting Diabetes
=======================================================================
Column {data-width=500}
-----------------------------------------------------------------------

### Information Collection
We need all 9 fields be filled out in order to make the best prediction.
Incomplete information might infer the model reliability. 


```{r}
sliderInput("Pregnancy", label = "Pregnancy:",
            min=0,max=17,value=0,step=1)
numericInput("BMI", "BMI:", 20, min = 0, max = 81, step=0.1)
numericInput("Age", "Age:", 20, min = 0, max = 120, step=1)
numericInput("Glucose", "Glucose:", 10, min = 0, max = 200, step=1)
numericInput("BloodPressure", "Blood Pressure:", 60, min = 0, max = 120, step=1)
numericInput("SkinThickness", "Skin Thickness:", 10, min = 0, max = 110, step=1)
numericInput("Insulin", "Insulin:", 10, min = 0, max = 744, step=1)
numericInput("PedigreePedigreeFunction", "Diabetes Pedigree Function:", 0.5, min = 0, max = 1, step=0.001)


```

Column {data-width=500}
-----------------------------------------------------------------------

### Results from SVM Model

```{r}
test=reactive({
Pregnancies=input$Pregnancy
Glucose=input$Glucose
BloodPressure=input$BloodPressure
SkinThickness=input$SkinThickness
BMI=input$BMI
Insulin=input$Insulin
DiabetesPedigreeFunction=input$PedigreePedigreeFunction
Age=input$Age
test=data.frame(Pregnancies,Glucose,BloodPressure,SkinThickness,BMI,Insulin,DiabetesPedigreeFunction,Age)})
```
#### According to our calculation, your chance (out of 1) of having diabetes is:
```{r}
renderPrint({
prediction_svm <- predict(svm_model,test(), probability = T)
pred_prob_svm <- attr(prediction_svm, "probabilities")[,"1"]
print(pred_prob_svm)
})
renderText({
prediction_svm <- predict(svm_model,test(), probability = T)
pred_prob_svm <- attr(prediction_svm, "probabilities")
if (prediction_svm ==1){
  print('You are likely having Diabetes.')
}else{
  print('You do not seem to have Diabetes comparing to other patients in our database.')}
})
```

### Results from RF Model
#### According to our calculation, your chance (out of 1) of having diabetes is:
```{r}

renderPrint({

prediction_rf <- predict(rf_classifier,test(), probability = T)
prediction_prob_rf <- predict(rf_classifier,test(),type="prob")[,"1"]
print(prediction_prob_rf)

})
renderText({

prediction_rf <- predict(rf_classifier,test(), probability = T)
prediction_prob_rf <- predict(rf_classifier,test(),type="prob")
if (prediction_rf ==1){
  
  ('You are likely having Diabetes.')}
else{
  ('You do not seem to have Diabetes comparing to other patients in our database.')}
})
```

Documentation
=======================================================================
